A space-time skew-$t$ model for threshold exceedances

### 2016-06-18 LaTeX/response.tex

We wish to thank the associate editor and two reviewers for their thorough reviews and constructive suggestions.
To address the concerns raised in this review we have carefully edited the manuscript to clarify issues regarding the skew-$t$ parameterization and other notational issues.
We have also added some new simulation results to explore more complicated max-stable processes, and a discussion of the properties of the spatiotemporal extension of our model.
We hope you agree that these changes have improved the manuscript.

\section*{Response to the comments of the Associate Editor}
\subsection*{Major comments}
\begin{enumerate}[1.]
\item The model in (1) has a non-standard skew distribution notation. As far as I could check, the parameter $\sigma^2$ should have an inverse gamma with parameters $a = b$, and not arbitrary and different $a$ and $b$. The equality is important in the derivations and it is one of the reasons we get $2a$ degrees of freedom. Indeed, some reserachers adopt $a/2$ and $a/2$ (rather than $a$ and $b$) such that they end up with a degrees of freedom. Can you provide a web-supplement proof that your more general model with $a \neq b$ is a skew-$t$ distribution with $2a$ degrees of freedom, or provide a reference for this result?\\

\begin{response}
  We have assumed that the errors $v(\bs)$ have unit variance, and $z \sim N(0, 1)$, so we require separate $a$ and $b$ so that $a$ can control the shape/degrees of freedom while $b$ controls the scale.
  Viewed this way, if $a = b$, then the degrees of freedom would determine the scale which is clearly too restrictive.
  To help clarify this, we have slightly modified the parameterization so that the degrees of freedom are $a$, and added additional clarification of the role of $b$ in \sref{sts:skewt}.
  We have also added \aref{sta:otherparams} to demonstrate our more general model with $a \neq b$ is a skew-$t$ distribution with $a$ degrees of freedom.
\end{response}

\item In the simulations, you adopted a prior distribution for the $a$ and $b$ parameters in $\sigma_t^2 (\bs) \sim IG(a, b)$ as $a \sim$ Gamma$(0.1, 0.1)$ and $b$ with a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$. The parameter $\sigma_t^2(\bs)$ is clearly a crucial parameter as it controls the scale. Why this choice? Trial and error? What have you used in the real dataset? Is this supposed to be a default choice?\\

\begin{response}
  In the original submission, the priors for $a$ and $b$ were inadvertently switched, so $a$ should have had a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$, and $b \sim$ Gamma$(0.1, 0.1)$.
  With the reparameterization given in the response to Major Comment 1, this actually means that the prior for $a$ is now a discrete distribution over a mesh from 0.2 to 20 with spacing $0.2$.
  As we now clarify in the specification of these priors, $a$ is the scale-free shape parameter, and so this prior should be uninformative in all applications.
  In contrast, $b$ is a precision parameter, so its prior should be adapted to the scale of the response.
  In the real data analysis, we use the same discrete prior as in the simulation study, and find that for all settings in the cross validation, the posterior mean of $a$ is less than 3.
\end{response}

@@

Anyway, you should make clear that trading a focus only on the extreme tail data you loose efficiency by decreasing the number of (real) data available. Censoring on purpose seems a rather drastic technique.\\

\begin{response}
  This is a key argument in extreme value analysis which we realize we inadvertently took for granted.
  This is a rendition of the classic bias variance trade-off.
  A low threshold reduces variance by increasing the amount of data used to estimate parameters but risks inducing bias if the model is misspecified.
  On the other hand, a high threshold mitigates potential bias because only the tails must be modeled.
  Therefore, by using a high threshold fewer precarious assumptions are required, but this increases variance by discarding data.
  For our fairly flexible skew-$t$ process, we found that a moderate threshold optimized this balance for both simulated and ozone data.
  \sref{sts:censoring} now elaborates on this trade-off.
\end{response}

\item The temporal dependence structure is induced by putting AR(1) models on some of the skew-$t$ parameters.
It is not clear what sort of temporal dependence that induces on the data.\\

\begin{response}
  Using simulated data, we have evidence to suggest that the model exhibits extremal dependence over time.
  We are unable to give a formal proof of this, but we have included simulated $\chi$ plots for temporal dependence in \aref{sta:temporal}.
\end{response}
\end{enumerate}

@@

  \begin{response}
    As we now describe at the end of the first paragraph of the introduction, two objectives for our data analysis are spatial prediction and mapping the probability of an extreme event.
    In the extremes context, spatial prediction means using the values of monitored stations to determine whether an unmonitored site experienced an extreme event.
    Mapping the probability of extreme events could either be the marginal probability of an event, such as ozone exceeding 75ppb, or a map of an upper quantile, such as the map of the 99th percentile in Figure \fref{stfig:ozoneq99}.
  \end{response}

  \item There is a lot of confusion about the ``threshold'', and what this means in various contexts. Sometimes you discuss this as fixed and specified by regulation (e.g. p.3 l.5) but other times you refer to the role of censoring as not allowing low-moderate values influence the fit (e.g. Section 3.1 or the stated goal of modeling ``spatial extremes'' in the first line of Section 3) even though you state that the constant threshold is not extreme everywhere (e.g. in the Abstract, and p.15 l.3 where in fact the highest threshold you use in the application is the 0.06 quantile at some sites, and so hardly the tail!). The role of the threshold needs to be much more clearly explained in the context of the application (fixed thresholds of interest), and modeling the extremes. \\

  \begin{response}
    Yes, we can see how this could lead to confusion.
    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in \eref{steq:Yt}.
    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below \eref{steq:Yt}.
    As mentioned in the response to the previous comment, we also want to compute probabilities of extreme events (e.g. ozone exceeding 75ppb).
    We now refer to this as the probability above a high \emph{level}, $L$.
    That is, we might use the data above threshold $T = 50$ppb in order to estimate the probability of the response exceeding level $L = 75$ppb.
    Of course, this is only reasonable if $L \ge T$, and the user is free to select the $T$ that gives the best estimate of this probability.
  \end{response}

@@

  The proof in Web Appendix C also needs some greater attention to detail. It looks to me as if using defintion (7) and for a fixed radius $h$ then $\bs_1$ and $\bs_2$ need not be in different partitions almost surely, although this may be the case as $h \rightarrow \infty$. Perhaps I am wrong, but either way it would be helpful to have the full proof spelt out to avoid any confusion. \\

  \begin{response}
    We have provided a more complete proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$ in \aref{sta:proofsamepartition}.
  \end{response}

  \item Section 3.3: As you note, there are several places where you could introduce temporal dependence. Are you able to analyze any theoretical properties of the suggested form, or even how this matches the data? \\

  \begin{response}
    Yes, as noted in the response to AE (major comment 5), we have evidence using simulated data that our temporal construction maintains extremal dependence.
    This now appears in \aref{sta:temporal}.
    Thank you both for the suggestion!
  \end{response}

@@

  \item Section 5: The simulation study needs more details. For example, when fitting datasets (1)--(4) to model (6) (max-stable), were the margins of this estimated or assumed fixed? Also you fit trend terms in some models when the mean is constant, and do not comment on the rationale for this or what happens to the estimates of these parameters (are $\beta_\text{lat}$ and $\beta_\text{long}$ close to 0 and $\beta_\text{int}$ close to 10?)

  Furthermore the range of models simulated from is rather limited; for (4) you could simulate from a range of different max-stable processes (even if you don’t fit these, it might be interesting to see how the skew-t model handles them). \\

  \begin{response}
    We have added clarification for these points in the description of the simulation study.
    The parameters in the marginal distribution for all methods were estimated in a fully-Bayesian manner using MCMC methods.
    The constant mean function was chosen somewhat arbitrarily, but permits a common threshold to apply to all sites.
    Finally, we have also added a 5th design to the simulation study using a Brown-Resnick process to generate data, and find that the skew-$t$ method performs similarly to the other setting.
    For this new setting, we find that our model performs similarly to the case in setting (4) with some improvement over the max-stable method for lower quantiles (0.90 -- 0.98), and the max-stable method tends to perform better at more extreme quantiles.
  \end{response}

@@

  \item p. 8 final line: the phrase ``use a copula'' could be omitted as it seems superfluous and may introduce unnecessary confusion. Giving adequate details on the transformation and dependence structure is sufficient. \\

  \begin{response}
    We have replaced ``use a copula'' with ``use a transformation of a Gaussian random variable'' above \eref{steq:sigmastar} and \eref{steq:zstar}.
  \end{response}

  \item p. 12 final line: you fit the correct model in some of these cases, so is it still difficult to tell whether the parameters converge even in this situation? \\

  \begin{response}
    As with any Bayesian mixture model, the labels of the partitions are arbitrary.
    That is, we could swap the labels of knots 1 and 2, and this would not affect the value of the posterior.
    It would, however, change the values of $z_1$ and $z_2$ making it difficult to study convergence for these parameters.
    However, since the quantities of interest (spatial predictions and probability of extreme events) are defined marginally over the partitioning, label-switching does not affect convergence for these quantities.
    We have bolstered this explanation and included coverage probabilities for $\lambda$ and the degrees of freedom for design (2) and method (2) (skew-$t$ with 1 knot) at the end of \sref{sts:simdesign}.
  \end{response}

@@

    \begin{response}
      Conceptually we can imagine that $Y(\bs)$ is a random field that is only observed at $n$ locations as described in and around \eref{steq:stazzalini}.
      This notation could be problematic if the data consisted of multiple measurements at some sites that gave different values, but this is not our case and so we elect to retain this streamlined notation which is used, for example, throughout \citet{Gelfand2010}.
    \end{response}

    \item Page 5, equation (3). You used the $h = ||\bs_1 - \bs_2||$ (above equation (4)) to denote the distance between two specific fixed points and then $h = ||\bs - \bt||$ to denote the distance between two generic locations. The latter (more general) should also be used in equation (3) to avoid confusion; \\

    \begin{response}
      Thank you for pointing this out.
      We now use $\bs_1$ and $\bs_2$ in \eref{steq:matern}, \eref{steq:chih}, and \eref{steq:chiskewt} as opposed to $\bs$ and $\bt$ to avoid confusion between space and time.
    \end{response}

    \item Page 7, below equation (7). The way you defined the partition is not clear to me. Suppose that we consider two spatial knots. Then, according to your definitions we have the two partitions:
    \begin{align*}
      P_k = \{s \in \calD : k = \argmin_i ||s - w_i||\}, \quad \text{with} \quad k= 1,2.
    \end{align*}
    The index of the partition is also used to denote the value of the radius. Does this mean that we also consider all the values $s \in \calD$ such that for a given $w_k$ we have that $\min ||s - w_k|| \le k$? In this simple example, is $\calD$ formed by two disjoint sets? \\

@@

  \item The section on the spatio-temporal extension is hard to follow and the technical details are not immediate to verify, in the present form. The presentation should be simplified. Is it not much easier that you define first the AR(1) time series and then you define appropriate transformations that link these to the functional parameters of the skew-$t$ process and you clearly show that the finite dimensional distribution of the resulting process including time and space is skew-$t$? \\

  \begin{response}
    We have heeded your advice and swapped the transformation equations (now equations \eref{steq:wstar} -- \eref{steq:sigmastar}) and the latent time series definitions (now equations \eref{steq:ts-wstar} -- \eref{steq:ts-sigmastar}).
  \end{response}

  \item Section 5 Simulation study. The current session should be shortened. Furthermore the section should be divided into two parts. Before showing the performances of your method in predicting exceedances, you should show by means of a simulation study the ability of your approach when estimating the model parameters. Since that you are considering also replications over time it should be possible to estimate reasonably well the slant parameters and the degrees of freedom. It may be the case that you have to consider larger samples. Have you tried to consider alternative parameterizations? See e.g. \citet{ArellanoValle2008}. You should investigate more seriously this aspect. \\

  \begin{response}
    In simulation design (2), we generate the data using a skew-$t$ distribution with a single knot, and in method (2), we fit the data using a skew-$t$ distribution with a single knot.
    We have provided additional clarification in \sref{sts:simdesign} that in this case, we do appear to correctly estimate the skewness parameters and degrees of freedom.
    We have also added clarification that the difficulty in assessing convergence is only relevant for models with multiple partitions (i.e. models 4 and 5) specifically for the parameters that change across partitions.
    Because the parameters in our model do appear to properly converge when the design matches the method, we did not consider a centered parameterization as in \citet{ArellanoValle2008}.
  \end{response}

@@

\begin{enumerate}[1.]
  \item Page 4, first paragraph of Section 2. The references \citet{Beranger2016}, \citet{Azzalini2014} page 129 and \citet{Azzalini2003}, should be added for a clearer explanation of the additive representation; \\

  \begin{response}
    We have added these references to clarify the representation.
  \end{response}

  \item Page 4, equation (1). You used the symbol ``$T$'' at least 3 times, e.g. in order to denote: transposed operation, a threshold (at page 6), a time index (at page 8). Different symbols should be used to avoid confusion; \\

  \begin{response}
    The revision uses $^\top$ for transposes and $n_t$ for the number of time points.
  \end{response}

### 2016-06-18 LaTeX/skewt_rev.tex
manuscript and plot updates

\begin{abstract}
To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified threshold.
In the case of ozone, the threshold for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
We present a new method based on the spatial skew-$t$ process.
Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
We also introduce a transformed AR(1) time-series to allow for temporal dependence.
Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances over a fixed threshold.
\end{abstract}

@@

\section{Introduction}\label{sts:intro}
% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
\fref{stfig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
We analyze these data with the goals of spatial prediction for unmonitored locations and to map the probability of extreme events.

@@

A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
First, because we are interested only in high values, we want to ``let the tail speak for itself''.
That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
Our inference method will only use data which exceed a threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.

@@

We propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
Our model is a threshold exceedance model for the multivariate skew-$t$ distribution for a fixed threshold.
In this setting, we describe the threshold as fixed because it is specified in advance by regulatory compliance.
This differs from the more traditional extremes literature where a threshold is selected to be the value beyond which an extremal model is appropriate for the data.
We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed threshold at an unobserved location.
\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}

@@

\subsection{Skew-$t$ process} \label{sts:skewt}
Let $Y(\bs)$ be a spatial process defined for spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
The spatial skew-$t$ process can be written
\begin{align}
  Y(\bs) = \bX(\bs)^\top \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{steq:fullmodel}
\end{align}
$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{steq:matern}
\end{align}
where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.

@@

\subsection{Censoring to focus on the tails} \label{sts:censoring}
As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
Let
\beq \label{steq:Yt}
  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
      Y(\bs) \quad & \delta(\bs) = 1 \\
      T & \delta(\bs) = 0
  \end{array} \right.
\eeq
be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ to large increases variance.
We impute the censored values as a step in the MCMC algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.

\subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
We handle the problem of long-range asymptotic dependence with a random partition.
As discussed in \sref{sts:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
The model becomes
\begin{align} \label{steq:partition}
  Y(\bs) &= \bX(\bs)^\top \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu_{PP}$ over spatial domain $\calD \in \calR^2$.
The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{steq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
In other words, $P_k$ is composed of all sites for which the closest knot is $\bw_k$.
For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in \sref{sts:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
However, the bivariate distribution for sites in different partitions is neither Gaussian nor skew-$t$ and does not have asymptotic dependence.

The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
\begin{align}
  \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) + I(k_1 \neq k_2) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) \label{steq:chiskewt}
\end{align}
where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in \eref{steq:chiskew-t}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs_1 - \bs_2||$.
Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
In \aref{sta:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, assuming that the knots follow a homogeneous Poisson process.
This implies that $\lim_{h \rightarrow \infty} \chi(h) = 0$.
In practice we fix $K$ at a finite value and use a uniform distribution for the knots $\bw_1, \ldots, \bw_K$.
In \fref{stfig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.

@@

\section{Simulation study}\label{sts:simstudy}
In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.

>>>>

\begin{abstract}
To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified level.
In the case of ozone, the level for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
We present a new method based on the spatial skew-$t$ process.
Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
We also introduce a transformed AR(1) time-series to allow for temporal dependence.
Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances of a high level.
\end{abstract}

@@

\section{Introduction}\label{sts:intro}
% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a high level determined by government regulation.
More specifically, we consider the case of compliance for ozone.
A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
\fref{stfig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
We analyze these data with the goals of spatial prediction for unmonitored locations and to map the probability of extreme events.

@@

A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
First, because we are interested only in high values, we want to ``let the tail speak for itself''.
That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
Our inference method will only use data which exceed a pre-selected threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.

@@

We propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
Our model is a threshold exceedance model for the multivariate skew-$t$ distribution using a fixed threshold.
In this setting, we describe the threshold as fixed because it is specified in advance by regulatory compliance.
This differs from the more traditional extremes literature where a threshold is selected to be the value beyond which an extremal model is appropriate for the data.
We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.

@@

\subsection{Skew-$t$ process} \label{sts:skewt}
Let $Y(\bs)$ be a spatial process defined for spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
The spatial skew-$t$ process can be written
\begin{align}
  Y(\bs) = \bX(\bs)^\top \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{steq:fullmodel}
\end{align}
$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
\begin{align}
  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{steq:matern}
\end{align}
where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.

@@

\subsection{Censoring to focus on the tails} \label{sts:censoring}
As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
Let
\beq \label{steq:Yt}
  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
      Y(\bs) \quad & \delta(\bs) = 1 \\
      T & \delta(\bs) = 0
  \end{array} \right.
\eeq
be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
As our goal is to predict exceedances above a high level $L$, we should select a value for $T \le L$.
For example, in predicting ozone exceedances, we might set $T = 50$ppb in order to predict exceedances of $L = 75$ppb.
Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ to large increases variance.
We impute the censored values as a step in the MCMC algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.

\subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
We handle the problem of long-range asymptotic dependence with a random partition.
As discussed in \sref{sts:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
The model becomes
\begin{align} \label{steq:partition}
  Y(\bs) &= \bX(\bs)^\top \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
\end{align}
To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu_{PP}$ over spatial domain $\calD \in \calR^2$.
The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
\begin{align} \label{steq:subregions}
  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
\end{align}
In other words, $P_k$ is composed of all sites for which the closest knot is $\bw_k$.
For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a / 2, b / 2)$.
So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in \sref{sts:spatialskew}.
Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
However, the bivariate distribution for sites in different partitions is neither Gaussian nor skew-$t$ and does not have asymptotic dependence.

The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
\begin{align}
  \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) + I(k_1 \neq k_2) \chi_{\text{Gaus}}(h) \nonumber \\
         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) \label{steq:chiskewt}
\end{align}
where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs_1 - \bs_2||$.
Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
In \aref{sta:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, assuming that the knots follow a homogeneous Poisson process.
This implies that $\lim_{h \rightarrow \infty} \chi(h) = 0$.
In practice we fix $K$ at a finite value and use a uniform distribution for the knots $\bw_1, \ldots, \bw_K$.
In \fref{stfig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.

@@

\section{Simulation study}\label{sts:simstudy}
In this section, we present the results from a simulation study to investigate how the number of partitions and the amount of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.
